Non-word repetition in Yélî Dnye

### 2020-09-07 manuscript.tex
incorporated changes to analysis of substitutions and deletions

Additionally, we looked up each phoneme in each target word in the Phoible database (Moran \& McCloy, 2019). We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number. * \textbf{mc} do you agree with this decision or shall we do the alternative?*

@@

This study was approved as part of a larger research effort by \textbf{({\textbf{???}}) add info from Middy's FAIR reply}. Participation was voluntary, with children being invited to come and participate. Regardless of how they performed, children were provided with a snack as compensation. Children who came up to participate but then refused were nonetheless provided with the snack.

@@

Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 29 vowels deleted, and 9 consonants.

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (52 oral target vowels produced as nasal vowels, 58 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (197 oral vowels produced with a different quality; 23 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 266) than vice versa (2). Simple consonants were mispronounced as other simple consonants quite frequently (128 simple consonants mispronounced as other simple ones, compared to 62 complex ones).

\textbf{probably turn this into a table; use proportion of correctly produced rather than number; use first attempts rather than the whole data set}

@@

** this is higher than any previous report! can we be sure of it?**

>>>>

Additionally, we looked up each phoneme in each target word in the Phoible database (Moran \& McCloy, 2019). We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number.

@@

This study was approved as part of a larger research effort by Marisa Casillas, with the research protocol having been approved by the IRB committee entitled The Ethics Committee of the Faculty of Social Sciences (Ethiek Commissie van de faculteit der Sociale Wetenschappen; ECSW) of Radboud University, Nijmegen, The Netherlands. The approval for work outside of the Netherlands came as an amendment to the general research line on language development research (original request: ECSW2017-3001-474 Manko-Rowland; amendment: ECSW-2018-041). Participation was voluntary, with children being invited to come and participate. Regardless of how they performed, children were provided with a snack as compensation. Children who came up to participate but then refused were nonetheless provided with the snack.

@@

\textbf{use proportion of correctly produced rather than number; DONE}

\textbf{a2m use first attempts rather than the whole data set: having second thoughts. To be clear: the argument for using first attempts only is that all of the other analyses are on kids' first attempts, rather than all of their attempts, so we should use just first attempts here because it may confuse readers \& it is more consistent. BUT there are fewer mistakes this way, which is Options 1 and 2 below has lots of zeros\ldots{} CF Option 3 further down}

\textbf{a2m probably turn the paragraph on substitutions into a table: having second thoughts, the table would look just like a para, no?}

\textbf{Option 1: just numbers, first attempts only }
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 19 vowels deleted, and 4 consonants.

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (0 oral target vowels produced as nasal vowels, 0 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (42 oral vowels produced with a different quality; 0 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 0) than vice versa (0). Simple consonants were mispronounced as other simple consonants quite frequently (62 simple consonants mispronounced as other simple ones, compared to 0 complex ones).

\textbf{Option 2: numbers and proportions, first attempts only}
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 19 vowels deleted (about 1.17\% of all vowel targets), and 4 consonants deleted (about 0.24\% of all consonant targets).

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (0 oral target vowels produced as nasal vowels or about 0\% out of all oral vowel targets, 0 nasal target vowels produced as oral vowels or about NA\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (42 oral vowels produced with a different quality or about 0.29\% out of all oral vowel targets; 0 nasal vowels produced with a different quality or about NA\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 0 times or about NA\% out of all complex consonant targets) than vice versa (0 times or about NA\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (62 simple consonants mispronounced as other simple ones or about 0.30\% out of all simple consonant targets, compared to 0 complex ones or about NA\% out of all complex consonant targets).

\textbf{Option 3: numbers and proportions, ALL attempts}
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 29 vowels deleted (about 0.55\% of all vowel targets), and 9 consonants deleted (about 0.26\% of all consonant targets).

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (52 oral target vowels produced as nasal vowels or about 2.01\% out of all oral vowel targets, 58 nasal target vowels produced as oral vowels or about 6.46\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (197 oral vowels produced with a different quality or about 3.20\% out of all oral vowel targets; 23 nasal vowels produced with a different quality or about 7.08\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 266 times or about 5.55\% out of all complex consonant targets) than vice versa (2 times about 5.55\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (128 simple consonants mispronounced as other simple ones or about 0.48\% out of all simple consonant targets, compared to 62 complex ones or about 2.79\% out of all complex consonant targets).

@@

\textbf{TO DISCUSS TOGETHER this is higher than any previous report! can we be sure of it?}
